### 2007-02-27 pubs/papers/hpdc-gmw-2007/paper.tex
revisions throughout
git-svn-id: file:///localdisk/subversion/inca/trunk/pubs@6582 7dba3f4a-8be6-0310-8b3b-b4fec25ea7f3

%
%  $Description: Author guidelines and sample document in LaTeX 2.09$ 
%
%  $Author: ienne $
%  $Date: 1995/09/15 15:20:59 $
%  $Revision: 1.4 $
%

\documentclass[times,10pt,twocolumn]{article} 
\usepackage{dblfloatfix}
\usepackage{epsfig}
\usepackage{latex8}
\usepackage{listings}
\usepackage{subfigure}
\usepackage{times}
\usepackage{url}

%------------------------------------------------------------------------- 
% take the % away on next line to produce the final camera-ready version 
% takes away page numbers
\pagestyle{empty}

%------------------------------------------------------------------------- 
\begin{document}

\title{User-level Grid Monitoring with Inca 2}

\author{Shava Smallen, Kate Ericson, Jim Hayes, Catherine Olschanowsky \\
San Diego Supercomputer Center\\ University of California, San Diego\\ 
9500 Gilman Drive, La Jolla, CA 92093-0505, USA\\ 
\{ssmallen,kericson,jhayes,cmills\}@sdsc.edu\\
}

\maketitle
\thispagestyle{empty}

%1.  Motivate and differentiate the Grid monitoring Inca does
%2. Describe Inca 2 design and its benefits
%3. Illustrate that Inca 2 design is mature and being used in production by
%several Grids
%4. Be different from Inca 1 paper

\begin{abstract}
The primary goal in the creation of Grids is to provide unified and coherent
access to distributed computing, data storage and analysis, instruments, and
other resources to advance scientific exploration.  Grids combine multiple
complex and interdependent systems that span several administrative domains.
This complexity poses challenges for both the administrators who build and
maintain the Grid resources and the scientists who use them.  While other Grid
monitoring tools provide system-level information on the utilization of Grid
resources, the Inca system provides user-level Grid monitoring with periodic,
automated user-level testing of the software and services required to support
Grid operation.  Inca can be used by Grid operators, system administrators,
and application users to identify, analyze, and troubleshoot user-level Grid
failures thereby improving Grid stability.  In this paper, we describe the new
features of our current Inca release, Inca 2.  We then describe the
architecture of the Inca 2 system, in addition to use cases that describe two
Inca 2 deployments in production environments.  
\end{abstract}

%------------------------------------------------------------------------- 
\Section{Introduction}
\label{intro}

% Standard "Grids are good" intro paragraph.  But are complex.  
Grid systems provide unified and coherent access to distributed computing,
data storage and analysis, instruments, and other resources.  These systems
require the careful coordination of software packages, services, and
configurations across multiple, heterogeneous resources.  The TeraGrid
project~\cite{teragrid}, for example, manages the coordination of software and
services by deploying and monitoring a common user environment across
distributed, heterogeneous resources. The organization of TeraGrid software
and services is referred to as the Coordinated TeraGrid Software and Services
(CTSS).  TeraGrid's software and services uniformity simplifies access to its
resources, which consist of more than 102 teraflops of compute capability and
more than 15 petabytes of online storage, all interconnected by a network that
can transfer a terabyte of data in under 10 minutes.

% Complex means they are difficult to provide and maintain. Describe the
% stakeholders.   Monitoring is needed.
Providing and maintaining a stable infrastructure for these complex 
Grid systems poses challenges for both administrators who build and
maintain Grid resources and scientists who use them.  First, system
administration is distributed across multiple administrative domains 
requiring a significant amount of coordination.  This is typically performed by
a group of Grid administrators or operators who consider local site policies and
resource heterogeneity
and decide when software (including updates and patches) should be deployed to
resources and how it should be configured.  
Each site's system administrators, who are not necessarily Grid
experts, are then responsible for providing the required Grid software
components and services for their resource and debugging any problems that
arise.  Well-defined requirements and good communication are required in this
model.  Otherwise, inconsistencies between the resources arise that either
inconvenience the user or prevent them from using a particular resource
altogether.  A second challenge is that failures will occur over
time in the Grid services due to network or system failures, software
misconfiguration, or software bugs.  In order to address these challenges,
Grid monitoring can be used to detect such problems, leading to a more
stable and dependable Grid infrastructure.

% What types of Grid monitoring are around and what the limitations are
One approach to monitoring Grids, used by tools such as
MonALISA~\cite{monalisa} and GridICE~\cite{gridice}, is to aggregate and
display data from existing cluster or system administrator monitoring tools
such as Ganglia~\cite{ganglia}, CluMon~\cite{clumon}, or
Nagios~\cite{nagios}.  This provides a centralized, systems-level view of Grid
resources where low-level host statistics and queue information can be
examined.  This type of monitoring information is useful for showing the
utilization of Grid resources, but it does not provide the type of
high-level monitoring needed to detect user problems, such as incompatible
software versions, within the Grid infrastructure.

% What user-level monitoring provides and why we think it is good
User-level Grid monitoring is Grid infrastructure testing and performance measurement 
from a generic, impartial user's perspective.  The goal of user-level monitoring is 
to detect and fix Grid infrastructure problems before users notice them -- 
user complaints should not be the first indication of Grid failures.  
In our view, the following guidelines define user-level Grid monitoring:

\begin{itemize}

\item In order to reflect regular user experiences, tests or performance
measurements of the Grid infrastructure should be done from a standard user
account.  It is important not to execute from a system administrator's account,
since it may have special privileges and use custom shell initialization files.

\item Also in order to reflect regular user experiences, tests should be
written and configured using information directly from user documentation
(e.g., hostnames, ports, pathnames, etc.).  This may not always be possible
when documentation and testing are done in parallel in a pre-production
environment, but they should be closely coordinated activities.

\item Since Grids are dynamic, user-level tests or performance measurements
should be automated and executed periodically.

\item User-level tests and performance measurements should be written and
maintained within an iterative refinement process.  An iterative process is
required since Grids are dynamic and software environments change over time as
packages are upgraded.  It is also difficult to write and deploy a test
correctly the first time because of portability issues and incomplete user
documentation.  For example, it may take some iterations to determine whether a
detected failure is the result of a faulty test, user document, or Grid
resource.

\item In order to determine overall Grid health, monitoring must be flexible
enough to include many types of user-level tests and performance measurements.
Typically all of the commands listed in the user documentation and/or at least
one basic test for each software component of the Grid infrastructure is
required.

\item Tests or performance measurements that interact with Grid services
require authentication and should be executed using a standard user GSI
credential that is mapped to the standard user account.

\item Tests or performance measurements should be executed locally on all Grid
resources when appropriate so that all Grid access points available to users
are verified.  Similarly, it is important to execute some tests all-to-all
in order to detect site-to-site configuration errors such as authentication
problems.

\item The configuration of tests or performance measurements should be managed
centrally in order to ensure consistent testing of the resources, rather than
by system administrators who may not use the same information that is
available to users.

\end{itemize}

% Inca 1~\cite{inca1} first addressed user-level Grid monitoring but had
% limitations.  We learned lessons from our TeraGrid deployment and
% developed Inca 2.

\noindent In 2003 SDSC, in partnership with TeraGrid, began developing Inca
1~\cite{inca1} to implement a user-level Grid monitoring system with the above
features and started using it to
verify that CTSS was deployed consistently across all TeraGrid resources and
to monitor its status.  At that time, the only available user-level Grid
monitoring tools were the NCSA TestGrid script~\cite{ncsa-test} and
Grid Integration Test Script~\cite{gits}, both of which ran a fixed number of Grid tests and formatted
results in HTML.  Although these tools were easy to install and produced
useful information, they showed the view of the Grid from a single
resource, lacked automation, and were not easily extensible.

The initial
version of Inca was implemented as a client-server architecture.  It provided
data collection, data storage (accessible from a Web services interface and with 
limited archiving capabilities), and data display though Web status
pages.  Inca 1 was first deployed to TeraGrid in mid 2003 and released in late
2003. After running Inca 1 for a year and a half on TeraGrid, we learned some
valuable lessons and designed Inca 2, our current release.  Inca 2
contains a number of substantial improvements over Inca 1 with respect to
security, installation and maintenance, and storage capabilities.  Inca 2 has
been running on TeraGrid since November 2006, and in February 2007 a production
version of the software was released.  Currently, seven other Grids in the
U.S., Europe, and Australia use Inca.

% This motivated our development of a new version of Inca.
In the next section, we describe the design goals and features of Inca 2. 
We then describe the Inca 2 architecture and how it can be used to provide
user-level Grid monitoring.  In Section~\ref{usecases}, we describe two 
uses of Inca 2 for Grid software environment verification and benchmarking.  
Finally, we describe some future work and summarize the paper.

%------------------------------------------------------------------------- 
\Section{Inca 2 Features}

% Implements user-level Grid monitoring and who it benefits
Inca 2 is a system that provides user-level monitoring of Grid functionality
and performance.  It was designed to be
general, flexible, scalable, and secure, in addition to being easy to deploy
and maintain.  Inca benefits Grid operators who oversee the day-to-day
operation of a Grid; system administrators who provide and manage resources;
and users who run applications on a Grid.  Besides meeting the guidelines 
described in Section~\ref{intro}, Inca 2 has additional features
listed below.  A '*' after the feature description indicates it is a new
feature of Inca 2; other items note improvements to Inca 1 features.  The 
Inca system:

\begin{enumerate}

\item Collects a wide variety of user-level monitoring results (e.g., simple
test data to more complex performance benchmark output).

\item Captures the context of a test or benchmark as it executes (e.g.,
executable name, inputs, source host, etc.) so that system administrators have
enough information to understand the result and can troubleshoot system
problems without having to know the internals of Inca.

\item Eases the process of writing tests or benchmarks and deploying them into
Inca installations.

\item Provides means for sharing tests and benchmarks between Inca users. *

\item Easily adapts to new resources and monitoring requirements in order to
facilitate maintenance of a running Inca deployment.

\item Stores and archives monitoring results (especially any error messages) 
in order to understand the behavior of a Grid over time. The results
are available through a flexible querying interface.

\item Securely manages short-term proxies for testing of Grid services using
MyProxy~\cite{myproxy}. *

\item Measures the system impact of tests and benchmarks executing on
the monitored resources in order to tune their execution frequency to
reduce the impact on the resources. *

\end{enumerate}

\noindent The following section describes how the Inca architecture implements
these features.

%------------------------------------------------------------------------- 
\Section{Inca 2 Architecture}

\begin{figure*}[tbp]
  \centering
  \mbox{\epsfig{file=arch.eps, width=.7\textwidth}}
  \caption{\label{arch_fig} Inca architecture.}
\end{figure*}

% overview description of components
Figure~\ref{arch_fig} shows the architecture of Inca 2.
Inca 2 is composed of 3 core
components (in highlighted box):  the agent, depot, and reporter manager.
The \textit{agent} and \textit{reporter managers} coordinate the execution of
tests and performance measurements on the Grid resources and the
\textit{depot} stores and archives the results.  The inputs to Inca 2 are one
or more \textit{reporter repositories} that contain user-level tests and
benchmarks, called \textit{reporters}, and a configuration file describing how
to execute them on the Grid resources.  This configuration is normally created
using an administration GUI tool
called \textit{incat} (Inca Administration Tool).  The output or results
collected from the resources are queried by the \textit{data consumer} and
displayed to users.  The following steps describe how an Inca administrator
would deploy user-level tests and/or performance measurements to their
resources.

\begin{enumerate}

\item Using the guidelines described in the introduction, the Inca
administrator either writes reporters to monitor their Grid or uses existing
reporters in a published repository.

\item The Inca administrator creates a deployment configuration file that
describes the user-level monitoring for their Grid using incat and submits it
to the agent.

\item The agent
  \begin{enumerate}
    \item fetches reporters from the reporter repository
    \item creates a reporter manager on each resource
    \item sends the reporters and instructions for executing them to each reporter manager.
  \end{enumerate}

\item Each reporter manager executes reporters according to its schedule and
sends results (reports) to the depot

\item Data consumers display collected data (reports) by querying the depot
\end{enumerate}

\noindent The following subsections describe the Inca components in more
detail, using the order of the steps above.

\SubSection{Reporters}

\lstset{
  basicstyle=\scriptsize\ttfamily, 
  frame=single,
  keywordstyle=\textbf, 
  identifierstyle=, 
  commentstyle=\scriptsize\ttfamily, 
  stringstyle=\ttfamily, 
  numbers=left, 
  numberstyle=\scriptsize,
  stepnumber=2,
  firstnumber=1,
  showstringspaces=false} 
\begin{figure}[tbp]
\lstset{language=Perl} 
\subfigure[]{\lstinputlisting{grid.globus.gramPing}}
\lstset{language=XML,keywordstyle={}} 
\subfigure[]{\lstinputlisting{grid.globus.gramPing.out}}
\caption{\label{pingReporter}(a) An Inca reporter that tests the user-level
availability of a Globus GRAM server and (b) sample output from a Globus GRAM 
reporter.}
\end{figure}

An Inca reporter is an executable program that tests or measures some aspect
of a system or installed software.   Reporter executables are designed to be
easy to produce and can be run outside of the Inca system (e.g., by a system
administrator).  A reporter should be written using the user-level
monitoring guidelines described in the introduction and could be a simple
Globus~\cite{globus} GRAM gatekeeper ping test (see Figure~\ref{pingReporter})
or a more complex Grid application benchmark~\cite{grasp}.  Reporters must
support certain command line options and produce an XML document or
\emph{report}
according to the Inca reporter schema.  The goal of the reporter schema is to
accommodate multiple types of data and to capture enough information about the
reporter execution to diagnose any detected failures.  The reporter schema
consists of the following elements:

\begin{itemize}
\item a set of header tags describing the context of the reporter
execution:  GMT timestamp, hostname, reporter name, reporter version, working
directory, reporter path, and arguments;
\item an optional set of debug or information log tags similar to log4j~\cite{log4j} output;
\item a body containing the results expressed as any XML sequence; and
\item an exit status tag
indicating whether the reporter was able to complete its test or measurement,
and an optional error message.  
\end{itemize} 
\noindent Because the body of a report can be any XML sequence, it enables
reporters to express a wide variety of information.  Today, we have four
standard body schemas to express software version information, software
functionality or service tests results, usage information, and performance
results.

An extensible set of Perl APIs is provided to handle much of the effort of
writing reporters.  Figure~\ref{pingReporter} shows (a) an Inca reporter that
pings a pre-WS Globus GRAM server and (b) its corresponding output when
executed on a local machine.  The reporter uses two of the Perl APIs:
Inca::Reporter::SimpleUnit and
Inca::Reporter::GridProxy (both subclasses of Inca::Reporter).  The Inca::Reporter API provides convenience
functions for handling the required input arguments
(lines~\ref{pingReporter}(a) 12-14 produces lines \ref{pingReporter}(b)
10-31) and printing the Inca-compliant header and exit status XML tags
(lines~\ref{pingReporter}(a) 3-10, 24 produces \ref{pingReporter}(b)
1-9, 43-46).  It also includes some informational and debug logging functions
similar to log4j.  The log output has been very useful to system
administrators, who can view a summary of the test or benchmark that the
reporter performed without reading the reporter source code.  Line
\ref{pingReporter}(a) 15 shows a special log function, loggedCommand, that
executes a system command (using a 30 second timeout) and logs to the 'system'
level of the log XML output (producing lines \ref{pingReporter}(b) 32-36).  The subclass
Inca::Reporter::SimpleUnit is the API responsible for printing the simplest
standard reporter body schema to report software functionality or service
tests results; it handles the printing of the small body XML
(lines~\ref{pingReporter}(b) 39-41).  Finally, the
Inca::Reporter::GridProxy dependency (line~\ref{pingReporter}(a) 11) adds a proxy
dependency to the reporter, telling the Inca system to download a
short-term proxy before running this reporter (this is discussed further in
Section~\ref{security}).  Most current reporters use the Perl APIs and consist
of fewer than 30 lines of code.

\SubSection{Reporter Repositories}

% What are repositories and their key features
Reporter repositories are collections that consist of reporters, required
packages and libraries (including the Perl APIs described in the previous
section), and a catalog file.  Repository contents are usually accessible using a URL
and designed to be shared across multiple Inca deployments.  The catalog file,
patterned after APT's Packages.gz~\cite{apt}, allows the user to indicate
reporter dependencies on CPAN modules, tar.gz packages, and other external
sources.  Inca automatically deploys dependencies into a
subdirectory of its installation (more in Section~\ref{agent}); this enables
reporters to be deployed to resources with minimumal effort.  Each package
specified in the catalog is versioned, allowing Inca to
automatically distribute reporter code updates such as bug fixes.  Because of
this, Inca administrators who maintain repositories need to ensure
that updates to reporters work properly before committing them to a repository.

% We have a lot of reporters ready to be picked 
The Inca team publishes a reporter repository that contains 143 reporters
developed and tested for TeraGrid.  Most reporters are either version
reporters that collect version information from a software package and
serve as a most basic functionality test, unit test reporters that run a more involved
functionality test, or performance benchmark reporters.  The unit and version
reporters cover software such as Grid middleware, compilers, math libraries,
data tools, and visualization tools.  Current performance reporters
measure data transfer and execute Grid benchmarks as described in
Section~\ref{grasp}.

\SubSection{Inca administration tool (incat)}

\begin{figure*}[tbp]
  \centering
  \mbox{\epsfig{file=incat.eps, width=\textwidth}}
  \caption{\label{incat_fig} Screenshot of incat that shows the reporters
  and their attributes available from the provided reporter repository.}
\end{figure*}

% Incat is interface Inca admin uses to configure reporters to resources
% introduce series here; why suites are important It is designed to make the
% effort of configuring and maintaining 100s series on resources easilty
The Inca administrative tool (incat) is a GUI implemented in Java Swing that an
Inca administrator uses to configure user-level Grid monitoring on their
resources.  Incat allows an Inca administrator to choose which resources to
monitor and which reporters to deploy to those resources.  The configuration
is stored in a XML file and sent to the agent, which handles the implementation
(as described in the next section).  The Inca administrator can reload the
configuration at any point and make updates.  Incat provides a number of
conveniences that enable a large number of results to be managed and collected
with a minimum effort.

To begin reporter configuration, the Inca administrator enters the location of
one more reporter repositories, and incat loads and displays the reporters
available from them, as shown in Figure~\ref{incat_fig}.
For each reporter, the Inca administrator can specify the resources to run on,
input arguments, the runtime environment, the frequency of execution, resource
limits, and email notifications.  The reports that a reporter produces on a
resource when executed with specific set of input arguments and runtime
environment are known as a \emph{report series} (or \emph{series} for short).
Thus, the grid.globus.gramPing reporter in Figure~\ref{pingReporter} would
produce one report series when executed with host 'tg-login.sdsc.edu' and
another report series when executed with 'dslogin.sdsc.edu'.  A set of related
series can be grouped into \emph{suites} and shared across other Inca
deployments, e.g.,  a Globus suite or a data transfer
performance suite.  Suites can also be useful in determining interoperability
among Grids or whether an application's requirements are being
fulfilled on a Grid.

% Resource information organized to make it easier to maintain
For each monitored resource, an Inca administrator selects an access method
(SSH or Globus) to access the resource and start monitoring there (described
in Sections~\ref{agent} and ~\ref{rm}).  Resources can also be grouped
together and referenced in the series configuration.  This facilitates the
process of executing the same series on multiple resources.  Furthermore,
attributes called \emph{macros} can be attached to resource groups for
reference in series configuration (i.e., inputs
and runtime environment).  For example, it is useful to have a macro called
'jobmanager' since the batch system is likely to differ from one machine to
the next and multiple tests will require its value.  A resource macro can also
have multiple values in which case the series will be executed on the resource
once for each macro value.  The
benefit of resource groups and macros is that when a reporter is executed on a
resource multiple times with different input arguments, it can be represented
succinctly.  Similarly, when a resource attribute is changed it only has to be
changed in one place for every series that references it to be updated.
Inca's initial configuration and maintenance is therefore easier and less error prone.

\SubSection{Agent}
\label{agent}

The Inca agent is a server that implements the configuration specified by the
Inca administrator.  When it receives a configuration file from the Inca
administrator, it will determine which reporters should be executed on each
resource by expanding the resource macros and groups; it then stores the
expanded configuration information in the depot.  The agent stages and
launches an Inca component, called a reporter manager (described in the
following subsection), on each resource using either SSH or Globus.  Once a
reporter manager contacts its agent, the agent transmits the reporters to
execute along with their dependencies and a suite that contains their series
configuration and schedule of execution.  As an Inca administrator makes
updates to the configuration, the agent will push out new reporters and/or
suite changes to each of the affected resources.

To further manage the Inca deployment, the agent regularly pings the reporter
managers on each resource in order to detect when a reporter manager has
failed so it can be restarted.  For example, if a resource goes down for a
planned maintenance, the agent will attempt to restart the reporter manager
(by default every hour) until it succeeds.  The agent also maintains a local
cache of reporters and their dependencies; it periodically checks the
repositories for updates (every 4 hours by default) and automatically
distributes them to the reporter managers if the latest version is desired.  
The agent is implemented in Java.

\SubSection{Reporter Manager}
\label{rm}

The Inca reporter manager is a lightweight process responsible for managing
the schedule and execution of Inca reporters on a single resource. The
reporter manager receives reporter updates and dependencies from the agent
along with requests for reporter scheduling changes.  Running under a regular
user account, the reporter manager executes reporters on-demand or using an
internal cron scheduler, and sends reports to the depot for archiving. The
reporter manager monitors reporter system usage using the system command
\texttt{ps} and enforces limits (e.g., wall clock time, CPU time, memory).
System usage information is sent to the depot with each report.

The reporter manager is implemented in Perl since Perl is available by default
on UNIX machines.  The cron scheduling is implemented using the CPAN Perl
module Schedule::Cron.

\SubSection{Depot}

The Inca depot server is responsible for storing configuration information and
the data produced by reporters. The depot uses a relational database via
Hibernate~\cite{hibernate} so that it can use a variety of databases.  The
depot provides full archiving of reporter output and structures its schema
around series and the minimization of redundant data, i.e., only stores
changes from one report to another in a report series.  The depot also
supports comparisons that are executed on a newly received report; 
% need Jim's help with this sentence
currently regular expression matching and greater than/less than comparisons are 
supported.  Optionally, an Inca administrator can select to receive
email notification when the result of the comparison changes.

Data can be queried using SQL queries. Predefined queries exist to return the
latest report instances of a suite, a single report instance, or a report
history.  A Web services interface is also available to provide
unauthenticated query access to data.  The depot is implemented in Java.

\SubSection{Data Consumer}

The Inca data consumer is a Web application that queries the Inca depot for
data and displays it in a user-friendly format.  The data consumer is packaged
with Jetty~\cite{jetty} so that Web pages can be served out of the box.  A set
of extensible JSP tags and pages query the Inca depot for data (returned as
XML)
and a set of XSL stylesheets format the data as HTML.  An Inca administrator
can customize the data display by either modifying the XSL stylesheets or
writing their own JSP tags and pages.

\SubSection{Security}
\label{security}

To reduce the security risk of Inca, all Inca components communicate with each
other using SSL by default.  Sensitive information such as passwords is
encrypted before being stored to disk on the agent.  To handle reporters that
require a valid proxy credential to execute, the reporter manager fetches a
short-term proxy from a MyProxy~\cite{myproxy} server (as specified by an Inca
administrator) and then destroys the proxy once the reporter has completed.
MyProxy authentication information is fetched from the agent each time a proxy
credential is needed and then cleared from memory.  Generating short-term 
proxies only when needed is more secure than maintaining long-term
credentials and is easier to manage.

\SubSection{Scalability}

% Should illustrate that we've thought about how Inca can scale
\begin{figure}[tbp]
  \centering
  \includegraphics[width=\columnwidth]{vo.eps}
  \caption{Coordinated Inca Deployments.}
  \label{tg_coord_fig}
\end{figure}

Inca was designed to distribute control and storage workload among
multiple Inca components if needed.  For example, one depot could forward
storage requests and queries to another depot to distribute its workload or to
maintain a mirror for fault tolerance.  Figure~\ref{tg_coord_fig} shows a more
complex scenario where two Inca deployments coordinate with each other.  For
example, application communities like GEON have their own Grid but also have
access to resources on TeraGrid.   Two separate Inca deployments can result in
duplicate testing. Instead, one Inca deployment can coordinate the execution
of a suite with another deployment.
Figure~\ref{tg_coord_fig} shows this type of coordinated deployment, where Grid B 
has its own resources but also has access to resources in Grid A.
When an Inca administrator submits a suite to Grid B, the Inca agent running
on Grid B will determine which reporters run on Grid A resources and forward
them as a suite request to Grid A's agent.  If Grid B has permission to submit
suites, Grid A's agent will execute the suite and configure its depot to
forwards results back to Grid B's depot.

%-------------------------------------------------------------------------

\Section{Use Cases}
\label{usecases}

\SubSection{TeraGrid}

% Description and Requirements - Describe CTSS
Requirements for user-level Grid monitoring on TeraGrid initially included the
validation and verification of CTSS on each resource.  As described in the
introduction, CTSS was designed to
allow users to more easily run Grid jobs across its distributed, heterogeneous
resources.  CTSS version 3 (CTSSv3) is available from all TeraGrid compute
resources and contains approximately 30 software packages that provide
Grid tools and services (e.g., Globus, GSI-SSH, MyProxy), data management
tools (e.g., SRB, HDF5), and applications tools (e.g., MPICH, GCC).

To provide user-level monitoring of CTSSv3, Inca was originally
deployed on TeraGrid in 2003 using Inca 1 and was transitioned to Inca 2 in
November 2006.  In addition to CTSSv3 monitoring, Inca is now also being used
to collect job usage information from Globus 2 GRAM logs and to detect expired
host and CA certificates and CA CRLs.

% Usage of Inca - Describe reporters, configuration, and deployment.  Figure.
A number of version and unit test reporters were developed to test the software packages in CTSS.
Currently, there are 56 such reporters executing an average of 109 series on
each of TeraGrid's 18 resources.  They include 48 all-to-all tests (GSI-SSH,
GRAM, and GridFTP), 20 Grid-related tests, 28 data-related tests, 30
application tool-related tests, and 2 security-related tests.

The TeraGrid
Inca configuration makes extensive use of macros and resource groups and
currently uses 80 series and 82 resource macros to manage the total of 1,928
series executing on TeraGrid resources.
Figure~\ref{tg_deploy_fig} illustrates TeraGrid's Inca 2 deployment
configuration.  The
agent, depot and consumer components are hosted at SDSC on 
sapa.sdsc.edu.  A reporter repository for TeraGrid is also hosted
at SDSC on inca.sdsc.edu.

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\columnwidth]{tg-deploy.eps}
  \caption{TeraGrid's Inca 2 Deployment.}
  \label{tg_deploy_fig}
\end{figure}

% Results - Consumer view of CTSS, security, etc.
TeraGrid currently has six status pages to display the Inca monitoring
information: a summary of CTSSv3 test failures, a detailed grid of CTSSv3 test
results, a page showing grid job
usage, results for security tests, results for secure MDS tests, and a list of
all running reporters.

Figure~\ref{status_page_fig} shows a portion of the TeraGrid detailed grid of
CTSSv3 test results
page.  CTSSv3 software packages are listed along the column and TeraGrid
resources (anonymized) in the header rows.  Test results for each package and
resource are shown in the table body rows.

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\columnwidth]{status-page.eps}
  \caption{Portion of the TeraGrid CTSSv3 detailed status page with the
  resources anonymized.}
  \label{status_page_fig}
\end{figure}

\SubSection{GrASP}
\label{grasp}

% Describe GrASP probes
The Grid Assessment Probes (GrASP)~\cite{grasp} are designed to 
serve as simple grid application kernel exemplars as well as a 
set of diagnostic tools. They test and measure the performance of 
basic Grid functions including file transfers, remote execution, 
and information services queries.  There are two basic probes: circle and 
gather.  The circle probe transfers a 100 MB file around a ring of
Grid nodes and measures the transfer time at each step.  The gather probe
transfers a 100 MB file from some number of source nodes to a compute node,
runs a computation through the job queue on the compute resource, and then
transfers a 100 MB file to a single destination node.  The transfer of the
first set of files is executed in parallel.  Transfer, compute, and queue times are recorded.

The GrASP probes yield data about Grid infrastructure performance and reliability.
In order to be most useful, probe data must be gathered 
repeatedly over time and archived.
In addition to storing the performance measurements,
any error messages encountered must be stored as well.    
In order to
gather historical data, the GrASP probes were modified
to follow the reporter specifications and deployed using Inca.
Porting the original code to follow the reporter specifications proved to be
trivial.  A number of scripts were then written to graph and analyze the GrASP
data and related error messages stored in the Inca depot.

\begin{figure}[tbp]
  \centering
  \mbox{\epsfig{file=errorHistogramTG.eps, width=\columnwidth}}
  \caption{\label{errorHist_fig} Number and types of errors collected by
  GrASP benchmark execution over a six-month period on TeraGrid.}
\end{figure}

Using Inca, GrASP was deployed on TeraGrid and GEON for a period of over 6
months~\cite{grasp2}.  During this deployment, the execution time of Preco, 
a Grid application run on the TeraGrid, was predicted within 12\% (excluding setup and cleanup time).
A recurring Globus 2 bug was also identified as shown in Figure ~\ref{errorHist_fig}.  The figure shows the frequency and types (shortened) of 
system errors that were detected while executing 
GrASP on TeraGrid in 2005. The highest occurring
error was number 9, Assertion GLOBUS FALSE, which was the result of a bug in
the Globus 2 GRAM server.

\SubSection{Other Uses}

Other Inca users include the European DEISA Grid and the UK National Grid
Service (NGS).  DEISA and NGS began using Inca in late 2004 and early 2005 respectively.  
Similar to TeraGrid, DEISA uses Inca to verify their Common
Production Environment (DCPE) across their 11 heterogeneous resources.  They
currently have 40 version reporters to verify each of the packages included in
DCPE.  In addition, they have unit reporters for the DEISA Applications Test
Suite (DATS) to check basic operation of the C/C++/Fortran compilers, the MPI
library, some numerical libraries, OpenMP jobs, CORBA middleware, and a third
party application.  DEISA is currently using Inca 1 and will transition to
Inca 2 in the next few months.

NGS currently has Inca 2 deployed on 8 resources to verify their Grid services 
and compilers.  They have wrapped the GITS~\cite{gits} tests using a new
API developed by the GITS developer.  The GITS reporters include 17 tests, 
such as small job submissions and data transfers, that verify Globus and GSI-SSH functionality.
NGS also has 12 tests for their compilers and
utilize 2 security reporters from the Inca repository.

%------------------------------------------------------------------------- 
\Section{Future Work}

In order to improve problem detection and diagnosis for system administrators,
we plan to add graph rendering and statistical
analysis capabilities to the Inca data consumer.  This will improve troubleshooting capabilities by
allowing errors to be understood in a historical context.  It will also enable
summary displays analogous to those generated by Web page statistics packages
like Webalizer~\cite{webalizer} that provide a number of useful "Top"
statistics.  We expect feedback from TeraGrid and other Inca users on the
exact types of statistics and graphs that will prove to be most useful.  Our
goal is to offer a default set of statistics and graphs that can be used out
of the box and provide the capability to add others.

Also to aid system administrator troubleshooting,
we plan to enhance Inca with a knowledge base that can be used to
suggest solutions to problems (or point to known bugs if there is no solution
available).  The knowledge base would be populated by solutions provided by
system administrators -- the content would be self-moderated in order to
improve its quality (by voting or adding comments to each other's
contributions).  A knowledge base could provide a potentially vital resource
for system administrators and offer a forum for sharing
invaluable experience and skills.

Finally, we plan to improve the fault tolerance of Inca by enhancing the
reporter manager so that it caches reports locally if it is unable to
contact the depot or sends reports to a backup depot if one is available.
The agent will similarly be enhanced to cache configuration changes locally or
send configuration to a backup depot if one is available in case of a depot
failure.

%------------------------------------------------------------------------- 
\Section{Summary}

~\newpage

%------------------------------------------------------------------------- 
\bibliographystyle{latex8}
\bibliography{paper}

\end{document}