### 2012-07-13 pubs/papers/inca-analysis/Inca-anaylsis.tex
git-svn-id: file:///localdisk/subversion/inca/trunk/pubs@14852 7dba3f4a-8be6-0310-8b3b-b4fec25ea7f3

\documentclass{sig-alternate}
\usepackage{subfigure,url}

\begin{document}

\title{}

\numberofauthors{1}

\author{
\alignauthor
Shava Smallen, Sameer Tilak, etc. \\
\affaddr{University of California, San Diego} \\ 
\affaddr{9500 Gilman Drive, La Jolla, CA 92093-0505, USA}\\ 
\email{\{ssmallen,sameer\}@sdsc.edu}\\
}

\maketitle

\begin{abstract}

\end{abstract}

\section{Introduction}

\section{Future Grid Overview}

\section{Inca Architecture}

\section{Use cases}

\section{Test suite classes}

\section{Experimental results}

\section{Future Work}

\section{Conclusion}

\bibliographystyle{abbrv}
\bibliography{wsn-dc}

 \end{document}